---
title: "P8105 Homework 1"
author: "By Wenda Yoanda, UNI:wy2486" 
output: github_document
---
```{r echo=FALSE, message=FALSE}
library(tidyverse)
```

## Problem 1

```{r echo=FALSE, message=FALSE}
data("penguins", package = "palmerpenguins")
```
**Dataset Description:**  
The `penguins` dataset contains information about penguins from 3 species: *Adelie*, *Gentoo*, and *Chinstrap*, observed on 3 different islands, namely *Torgersen*, *Biscoe*, and *Dream*. There are 8 important variables in this dataset that describe each penguin, including *species, island, bill length (mm), bill depth (mm), flipper length (mm), body mass (g), sex,* and *year*.

**Size of the Dataset:**  
There are `r nrow(penguins)` observations and `r ncol(penguins)` variables.

**Mean Flipper Length:**  
The mean flipper length is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2)` mm.
<!--Calculates the mean flipper length, ignores missing values (NA), and rounds the result to 2 decimal places.
ChatGPT was used to help troubleshoot why the mean returned NA and to understand the use of na.rm = TRUE.-->

**Scatterplot of *flipper_length_mm (y)* vs *bill_length_mm (x)*:**
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(penguins, aes(
  x = bill_length_mm,
  y = flipper_length_mm,
  color = species
)) +
  geom_point()

ggsave("penguins_scatterplot.png")
```

## Problem 2

**Creating a New Dataframe -- Flipper Category:**

```{r echo=FALSE, message=FALSE, warning=FALSE}
set.seed(1)

p2flippercat_df = tibble(
  random_sample = rnorm(10),
  logical_vector = random_sample > 0,
  character_vector = c(
    "Penguin_A", "Penguin_B", "Penguin_C", "Penguin_D", "Penguin_E",
    "Penguin_F", "Penguin_G", "Penguin_H", "Penguin_I", "Penguin_J"
  ),
  factor_vector = factor(c(
    "short", "medium", "long", "short", "medium",
    "long", "short", "medium", "long", "short"
  ))
)

p2flippercat_df
```
**Pulling Mean of Each Variable in The Dataframe:**

* The mean of random_sample is `r round(mean(pull(p2flippercat_df, random_sample)), 2)`  
* The mean of logical_vector is `r round(mean(pull(p2flippercat_df, logical_vector)), 2)`  
* The mean of character_vector is `r round(mean(pull(p2flippercat_df, character_vector)), 2)`  
* The mean of factor_vector is `r round(mean(pull(p2flippercat_df, factor_vector)), 2)`  

<!--Calculates the mean of each variable, pull and rounds the result to 2 decimal places.-->

As shown above, calculating the mean of the character vector and factor vector produces *NA* (does not work), while it works for the random sample and logical vector. The random sample is numeric, so its mean can be calculated directly. For the logical vector, R treats *TRUE as 1* and *FALSE as 0*, so the mean can also be calculated. The character and factor vectors are not numeric, so their means cannot be calculated directly.

**Converting Variables From One Type To Another:**

```{r results="hide", warning=FALSE}
as.numeric(pull(p2flippercat_df, logical_vector))

as.numeric(pull(p2flippercat_df, character_vector))

as.numeric(pull(p2flippercat_df, factor_vector))
```

When I ran the code chunk above, I found that:

* For logical_vector, R converts *TRUE* to 1 and *FALSE* to 0. 
* For character_vector, R cannot convert values like *"Penguin_A", "Penguin_B", "Penguin_C", etc* into numbers, so the results are NA. 
* For factor_vector, R converts the categories into numbers, where *"long"* = 1, *"medium"* = 2, and *"short"* = 3. However, these numbers are just codes for the different categories and do not have actual numerical meaning.

This helps explain the results when calculating the mean. The mean works for the logical vector because R can treat *TRUE* and *FALSE* as 1 and 0. However, a mean does not make sense for the character or factor vectors because they do not contain actual numerical measurements.



