---
title: "P8105 Homework 1"
author: "By Wenda Yoanda, UNI:wy2486" 
output: github_document
---
```{r echo=FALSE, message=FALSE}
library(tidyverse)
```

## Problem 1

```{r echo=FALSE, message=FALSE}
data("penguins", package = "palmerpenguins")
```
**Dataset Description:**  
The `penguins` dataset contains information about penguins from 3 species: *Adelie*, *Gentoo*, and *Chinstrap*, observed on 3 different islands, namely *Torgersen*, *Biscoe*, and *Dream*. There are 8 important variables in this dataset that describe each penguin, including *species, island, bill length (mm), bill depth (mm), flipper length (mm), body mass (g), sex,* and *year*.

**Size of the Dataset:**  
There are `r nrow(penguins)` observations and `r ncol(penguins)` variables.

**Mean Flipper Length:**  
The mean flipper length is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2)` mm.
<!--Calculates the mean flipper length, ignores missing values (NA), and rounds the result to 2 decimal places.
ChatGPT was used to help troubleshoot why the mean returned NA and to understand the use of na.rm = TRUE.-->


